---
title: 'Case Study: Adventure Works'
author: "Adejumo Ridwan Suleiman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

# Install Packages
```{r}
#install remotes
#install.packages("remotes")

#install and load adventure works database
remotes::install_github("adejumoridwan/adventureWorks")
library(adventureWorks)

#install tidyverse
library(tidyverse)

```

# Data in the database
- sales
- customers
- products
- budget
- calendar
- territory

# Total Sales Made in each product sub category
```{r}
inner_join(x = sales, 
           y = products,
           by = "ProductKey") |> 
  group_by(Category,SubCategory) |> 
  summarize(Total_Sales = round(sum(SalesAmount))) |> 
  arrange(desc(Total_Sales))

```

# Profit Margin made in each Product category
```{r}
inner_join(x = sales, 
           y = products,
           by = "ProductKey") |> 
  mutate(Profit_Margin = 
           (SalesAmount - ProductStandardCost)/sum(SalesAmount)*100)|> 
  group_by(Category,SubCategory) |> 
  summarize(Total_Profit_Margin = sum(Profit_Margin)) |> 
  arrange(desc(Total_Profit_Margin)) |> 
  ggplot(aes(x = Total_Profit_Margin,
             y = reorder(SubCategory, Total_Profit_Margin),
             fill = Category)) +
  geom_col() +
  labs(x = "Profit Margin",
       y = "Products Sub Category",
       title = "Profit Margin for each Product Subcategory")
```

# Profit Margin of Sales over time
```{r}
sales |> 
  mutate(Profit_Margin = 
           (SalesAmount - ProductStandardCost)/sum(SalesAmount)*100,
         OrderDate = lubridate::dmy(OrderDate)) |> 
  group_by(OrderDate) |> 
  summarize(Profit_Margin = sum(Profit_Margin)) |> 
  ggplot(aes(x = OrderDate,
             y = Profit_Margin)) +
  geom_line() +
  labs(x = "Order Date",
       y = "Profit Margin",
       title = "Profit Margin of Sales over time")
```

# What happened in 2019
```{r}
sales |> 
  mutate(Profit_Margin = 
           (SalesAmount - ProductStandardCost)/sum(SalesAmount)*100,
         OrderDate = lubridate::dmy(OrderDate)) |> 
  group_by(OrderDate) |> 
  summarize(Profit_Margin = sum(Profit_Margin)) |> 
  filter(OrderDate >= "2019-01-01") |> 
  ggplot(aes(x = OrderDate,
             y = Profit_Margin)) +
  geom_line() +
  labs(x = "Order Date",
       y = "Profit Margin",
       title = "Profit Margin of Sales over time")
```

# What happened after July
```{r}
inner_join(x = sales, 
           y = products,
           by = "ProductKey") |>  
  mutate(Profit_Margin = 
           (SalesAmount - ProductStandardCost)/sum(SalesAmount)*100,
         OrderDate = lubridate::dmy(OrderDate)) |> 
  filter(OrderDate >= "2019-07-01") |> 
  group_by(Category) |> 
  summarize(Profit_Margin = sum(Profit_Margin))
```
So products belonging to bike categories were not sold, no wonder sales dropped

# Customers where most profit were made from
```{r}
inner_join(x = sales, 
           y = customers,
           by = "CustomerKey") |> 
  mutate(Profit_Margin = 
           (SalesAmount - ProductStandardCost)/sum(SalesAmount)*100) |> 
  group_by(Occupation, MaritalStatus, Gender) |> 
  summarize(Total_Profit = sum(Profit_Margin)) |>
  arrange(desc(Total_Profit))
```

# Exercises

# Question 1
Which sub category was sales made most from for single female.  
1. Touring Bikes
2. Mountain Bikes
3. Helmets
4. None of the Above - Ans
```{r}
inner_join(x = sales, 
           y = customers,
           by = "CustomerKey") |> 
  inner_join(y = products,
              by = "ProductKey") |> 
  filter(Occupation == "Professional",
         MaritalStatus == "S",
         Gender == "F") |> 
  group_by(Category, SubCategory) |> 
  summarize(Total_Sales = sum(ExtendedAmount)) |> 
  arrange(desc(Total_Sales))
```

## Question 2
Calculate the total tax on the product category Bikes.  
1. 2,265,451.62 - Ans
2. 56,060.80
3. 27,181.81
4. None of the Above

```{r}
inner_join(x = sales, 
           y = products,
           by = "ProductKey") |> 
  group_by(Category) |> 
  summarize(Total_Tax = sum(TaxAmt)) |> 
  arrange(desc(Total_Tax))
```

## Question 3
What is the total profit on bikes made from customers who are Professionals.
Note: Profit = Selling Price - Cost Price.  
1. 3,916,901  
2. 3,912,902
3. 3,716,902
4. None of the Above - Ans
```{r}
inner_join(x = sales, 
           y = customers,
           by = "CustomerKey") |> 
  inner_join(y = products,
              by = "ProductKey") |>
  mutate(Profit = SalesAmount - ProductStandardCost) |> 
  filter(Occupation == "Professional") |> 
  group_by(Category) |> 
  summarize(Profit = sum(Profit)) |> 
  filter(Category == "Bikes")
```
```{r}
library(ggplot2)
diamonds |> 
  ggplot(mapping = aes(x = cut)) + 
  geom_bar() +
  labs(title = "My Plot")
```



