---
title: "Date Tidying with tidyr"
author: "Adejumo Ridwan Suleiman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

# Installing and loading tidyr
```{r}
#install.package("tidyr)
library(tidyr)
```

# Reshaping with Pivot longer
```{r}
head(relig_income)
```


```{r}
relig_income %>%
  pivot_longer(!religion, 
               names_to = "income", 
               values_to = "count")
```

# Reshaping with Pivot wider
```{r}
head(us_rent_income)
```

```{r}
us_rent_income %>%
  pivot_wider(
    names_from = variable,
    values_from = estimate,
    values_fill = 0
  )
```

#Separating
```{r}
head(table3)
```

```{r}
table3 %>% 
  separate(col = rate, 
           into = c("cases", "population"),
           sep = "/")
```


#Uniting
```{r}
head(table5)
```


```{r}
table5 %>% 
  unite(col = new, 
        century, year)
```

# Exercises

## Question 1
Which of the follwing is an argument in the function `pivot_longer`?  
1. names_to - ANS  
2. value_to  
3. names_from  
4. values_from  
5. name_to  

## Question 2
Which of the following is an argument in the function `pivot_wider`?    
1. names_to  
2. value_to  
3. names_from - ANS  
4. value_from  
5. name_to   

## Qustion 3
Which of the following arguments will fill missing values after using a `pivot_wider` on a data set?  
1. value_fil  
2. values_fill - ANS  
3. fill  
4. fill_na  
5. na_fill  