---
title: "Data Manipulation With Dplyr"
author: "Adejumo Ridwan Suleiman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

# Installing dplyr
```{r}
#install.packages("dplyr")
library(dplyr)
```


# Installing nycflights13
This data is an RDMS: a Relational Database Management System, it is made up of more than one table of data which are related to each other.
- flights
- airlines
- airport
- planes
- weather

```{r}
#install.packages("nycflights13)
library(nycflights13)
```


## Flights (Main Data)
Details of all flights in the year 2013
```{r}
head(flights)
```


## Airlines Names
```{r}
head(airlines)
```

## Airport Metadata
```{r}
head(airports)
```

## Planes Metadata
```{r}
head(planes)
```

## Weather (hourly)
```{r}
head(weather)
?weather
```

#  Grouping and Summarizing

```{r}
month_delay <- flights |> 
  group_by(month) |> 
  summarize(avg_dep_delay = mean(dep_delay, na.rm = TRUE),
            avg_arr_delay = mean(arr_delay, na.rm = TRUE))
```

```{r}
carrier_delay <- flights |> 
  group_by(carrier) |> 
  summarize(avg_carr_dep_delay = mean(dep_delay, na.rm = TRUE),
            avg_carr_arr_delay = mean(arr_delay, na.rm = TRUE))
```

# Arranging
```{r}
carrier_delay |> 
  arrange(desc(avg_carr_dep_delay))
```

# Filtering
```{r}
flights |> 
  filter(month == 1 & dep_delay < 0 & arr_delay < 0)

carrier_delay
```

# Selecting
```{r}
flights |> 
  select(!(hour:time_hour))
```

# Creating Variables
```{r}
carrier_speed <- flights |> 
  mutate(speed = distance/(air_time/60)) |> 
  select(carrier, speed) |> 
  group_by(carrier) |> 
  summarize(avg_speed = mean(speed, na.rm = TRUE)) |> 
  arrange(desc(avg_speed))
```


```{r}
flights |> 
  mutate(speed = distance/(air_time/60), .keep = "all")
```


# Renaming
```{r}
flights |> 
  rename(destination = dest)
```

# Mutating Joins
```{r}
top5_carrier_speed <- carrier_speed |> 
  head(5)
top5_carrier_speed
```

```{r}
full_join(x = top5_carrier_speed, 
           y = airlines, 
           by = "carrier")

inner_join(x = carrier_delay, 
           y = airlines, 
           by = "carrier") |> 
  select(carrier, name, avg_carr_dep_delay, avg_carr_arr_delay) |> 
  arrange(avg_carr_dep_delay) |> 
  head(5)
```

```{r}
inner_join(x = top5_carrier_speed, y = airlines, by = "carrier")
```

# Filtering Join
```{r}
anti_join(x = airlines,
          y = top5_carrier_speed,
          by = "carrier")
```

# Exercises and Solutions

## Question 1
Using the flights data set, which carrier have the highest average speed.
Note: Remember to set na.rm = TRUE when calculating the average speed.
1. Hawaiian Airlines Inc. - ANSWER
2. Virgin America
3. Alaska Airlines Inc.
4. Frontier Airlines Inc.
5. United Air Lines Inc.
```{r}
avg_speed_table <- flights |> 
  mutate(speed = distance/air_time*60) |> 
  group_by(carrier) |> 
  summarize(avg_speed = mean(speed, na.rm = TRUE)) |> 
  arrange(desc(avg_speed))

avg_speed_table

inner_join(x = avg_speed_table, 
           y = airlines, 
           by = "carrier")
```

## Question 2
How many flights in the month of December had no departure and  arrrival delay.
1.39
2.37
3.32
4.31 ANSWER
5.34
```{r}
flights |> 
  filter(arr_delay == 0 & dep_delay == 0 & month == 12)
```

## Question 3
What is the distance covered in Kilometers for the flight with id number 4646 and tailnum N273WN. Note: A mile is 1.6 Kilometers
1. 115.625 Kilometers
2. 296 Kilometers - ANSWER
3. 290 Kilometers
4. 78 Kilometers
5. 234 Kilometers
```{r}
flight_4646_N273WN <- flights |> 
  filter(flight == 4646 & tailnum == "N273WN") |> 
  mutate(dist_kil = distance*1.6)

flight_4646_N273WN |> 
  select(dist_kil)
```

## Question 4
The manufacturer of the plane in Question 3 is:
1. SIKORSKY
2. EMBRAER
3. AIRBUS
4. BOEING - ANSWER
5. GULFSTREAM AEROSPACE
```{r}
inner_join(x = flight_4646_N273WN,
           y = planes, 
           by = "tailnum") |> 
  select(manufacturer)
```
## Question 5
What is the tailnum of the fastest Air Plane departing New York.
1. N819AW
2. N382HA
3. N654UA
4. N228UA - ANSWER
5. N315AS

```{r}
plane_speed <- flights |> 
  mutate(speed = distance/(air_time/60)) |> 
  select(tailnum, speed) |> 
  group_by(tailnum) |> 
  summarize(avg_speed = mean(speed, na.rm = TRUE)) |> 
  arrange(desc(avg_speed)) |> 
  head(5)

plane_speed
```

```{r}
inner_join(x = plane_speed,
           y = planes,
           by = "tailnum")
```

