---
title: "Data Visualization With ggplot2"
author: "Adejumo Ridwan Suleiman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

# Installing and Loading ggplot2
```{r}
#install.packages("ggplot2")
library(ggplot2)
```

# Diamonds dataset
```{r}
head(diamonds)
```
# Economics dataset
```{r}
head(economics)
```

# Barplots

## Vertical Bar Plot
```{r}
# Most occurring diamonds (change fill with "fill = ", change line colour with "colour = ")
diamonds |> 
  ggplot(mapping = aes(x = cut)) +
  geom_bar()
```

## Horizontal Bar Plot
```{r}
diamonds |> 
  ggplot(mapping = aes(y = cut)) +
  geom_bar()
```

## Stacked BarPlot
```{r}
diamonds |> 
  ggplot(mapping = aes(x = cut, fill = clarity)) +
  geom_bar()
```
## Grouped Barplot
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```

# Boxplots

## Comparing diamond cut and price
```{r}
diamonds |> 
  ggplot(mapping = aes(x = cut, y = price, colour = clarity)) +
  geom_boxplot()
```

# Histogram
```{r}
#Check for the distribution of diamond carat
diamonds |> 
  ggplot(aes(x = carat)) +
  geom_histogram()
```

# Scatter Plot
```{r}
# Check for the relationship between diamond price and carat
diamonds |> 
  ggplot(aes(x = price, y = carat)) +
  geom_point()

cor(diamonds$cut, diamonds$color, method = "spearman")
```

# Line Graph
```{r}
economics |> 
  ggplot(aes(x = date, y = pop)) +
  geom_line()
```


# Labelling
```{r}
diamonds |> 
  ggplot(mapping = aes(x = cut, y = price, colour = cut)) +
  geom_boxplot() +
  labs(x = "Cut", y = "Price", title = "Diamond Cut against Price", )
```

# Facetting
```{r}
diamonds |> 
  ggplot(mapping = aes(y = cut, fill = cut)) +
  geom_bar() +
  facet_grid(color ~ .)
```


# Exercises and Solutions

## Question 1
Using a boxplot to visualize diamond cut against depth percentage. Which diamond cut have the highest median depth percentage.
1. Fair - ANSWER
2. Good
3. Very Good
4. Premium
5. Ideal
```{r}
diamonds |> 
  ggplot(mapping = aes(x = cut, y = depth, color = cut)) +
  geom_boxplot()
```

## Question 2
Visualize using a line graph the change in United States Personal consumption expenditures in the economics data. In which year did the United States have the low personal consumption expenditure.
1. 1967 - ANSWER
2. 2000
3. 1999
4. 2015
5. 1990

```{r}
economics |> 
  ggplot(mapping = aes(x = date, y = pce)) +
  geom_line()
```

## Question 3
Using a suitable bin width, visualize the distribution of diamond carat in the diamonds data set. In which range of carat value do most diamonds fall in.
1. 0.5 - 1.5 - ANSWER
2. 4.0 - 5.0
3. 2.0 - 3.5
4. 0.0 - 0.5
```{r}
diamonds |> 
  ggplot(mapping = aes(x = carat)) +
  geom_histogram()
```

# Question 4
Which of the following can we use to plot a numerical variable against a numerical variable
1. Box Plot
2. Scatter Plot - ANSWER
3. Bar Plot
4. Line Plot
5. None of the Above

# Question 5
Which of the following codes in the options, will give the graph below.
SOLUTION - ALL THE GIVEN CODES WILL GIVE THE SAME GRAPH
```{r}
diamonds |> 
  ggplot(mapping = aes(x = cut)) +
  geom_bar()
```

```{r}
diamonds |> 
  ggplot() +
  geom_bar(mapping = aes(x = cut))
```

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut)) +
  geom_bar()
```


